### 2017-03-31 paper/oopsla17-submission/overview.tex
more overview

\paragraph{\textbf{Challenge (1): How to represent programs?}}

\paragraph{\textbf{Solution: Use a set of feature vectors}}

>>>>

\paragraph{\textbf{Challenge (1): How to represent programs?}}
The first and prime challenge we address is how to translate programs
into a format that machine learning techniques can understand.
%
Programs can be represented in many ways, from the raw text, to richly
structured abstract syntax trees and control-flow graphs, but none
of these representations are immediately amenable to machine learning.
%
Machine learning techniques typically take as input \emph{feature
vectors}, fixed-length vectors of real numbers where each element
represents some property of the input.
%
Thus, in order to apply machine learning to our problem we must
determine:
%
(1) what are interesting features of a program, and
%
(2) how can we represent a variable-size program as a fixed-length vector?

\paragraph{\textbf{Solution: Vectorized Abstract Syntax Trees}}
We start by assuming that we will extract features from the program's
abstract syntax tree (AST), a natural choice since the type checker
operates on ASTs.
%
An obvious set of interesting features is then the syntactic class of
each expression, \ie the classifier should be able to distinguish
between the |[]| and |+| expressions in \autoref{fig:sumList} because
they are formed by different constructors in the AST.
%
A programming language has a fixed set of AST elements, which suggests a
natural way to obtain fixed-size feature vectors: rather than translating
a program in a single feature vector, we translate it into a \emph{set} of
feature vectors, one for each expression in the program.
%
Thus, the |[]| and |+| expressions will each be given their own feature
vector, and we will ask the classifier to predict for each
\emph{independently} whether it should be blamed.

Of course, each expression occurs in some \emph{context}, and we would
like the classifier to be able make decisions based on the context as
well.
%
For example, we might expect that the recursive call |sumList t| is
unlikely to be at fault because |sumList| has a function type --- the
user may have called the wrong function or supplied the wrong argument,
but the application itself is probably correct.
%
Conversely, if the student wrote |h sumList t| (\ie forgot the |+|), we
might wish to blame the application rather than |h| because |h|
\emph{does not} have a function type.
%
Thus we \emph{extend} each expression's feature vector with
\emph{contextual features} of its parent and children.
%
These contextual features include the syntactic class of the neighboring
expressions, their inferred types, and, crucially, whether the
expression occurs in a minimal type error \emph{slice}.
%
A minimal type error slice of a program includes all expressions that
are necessary for the error to manifest and no more, this allows us to
tell the classifier which expressions could potentially be blamed ---
a change to an expression outside of the minimal slice cannot possibly
fix the type error.
